---
title: "COVID Vaccination Intention Among Adults in the United States - Final Project"
author: "Megan Richard"
date: "12/10/2020"
output:
  html_document: default
---

# Description of Research and Data
Due to the high prevalence of COVID-19, scientists are working hard to quickly find a safe and effective vaccine. Herd immunity will need to be achieved in order to stop COVID-19 from continuing to spread, and a vaccine will be imperative in achieving this goal. Understanding someone’s motivations, barriers, and other associated factors for getting vaccinated are important in increasing overall vaccination rates. In order to increase this understanding, this research study seeks to compare constructs of the health belief model, general demographic variables, and psychosocial variables with intention of COVID-19 vaccination among adults in the United States. The data was collected between October 2nd - October 10th of 2020 through a Google Survey that was disseminated on social media and among contacts of the survey author (me). The survey questions were carefully chosen by evaluating similar existing literature and cross-analyzing that with the information that is most critical in answering the research question. There are 524 observations and 20 variables total on the survey total. For the purpose of the data analysis, 8 variable will be looked at in more detail with 521 observations. (3 responses were dropped for analysis purposes because they identified as people who live outside of the United States.) Respondents with missing values will not be removed from the analysis in order to keep the sample size as large as possible.

## Research Question
What factors are associated with COVID-19 vaccine intention of adults in the United States? 

#### Outcome Variable 
Intention to get the COVID-19 Vaccine - Categorical 

#### Predictors 
* Political Beliefs - Categorical
* Highest Education Level - Categorical
* Flu shot in the past 12 months - Categorical
* Perceived Susceptibility - Categorical
* Perceived Severity - Categorical 
* Perceived Benefits - Categorical 
* Perceived Barriers - Categorical 


```{r include=FALSE , warning=FALSE, message=FALSE}

#Open Tidyverse 
library(package = "tidyverse") 

# Import the data 
covid <- read.csv(file = "COVID-19 Vaccine Survey Responses.csv")

#take out international - does not align with research question
covid.dirty <- covid %>% 
  select( Once.it.is.available..do.you.intend.to.get.the.vaccine.for.COVID.19.,
          Have.you.received.a.flu.shot.in.the.past.12.months., 
          What.is.the.highest.degree.or.level.of.school.you.have.completed., 
          Which.most.closely.aligns.with.your.political.beliefs.,
          I.am.at.risk.for.contracting.COVID.19.,
          COVID.19.can.be.a.serious.disease.,
          The.COVID.19.vaccine.will.prevent.me.from.contracting.COVID.19., 
          There.are.many.risks.associated.with.getting.the.COVID.19.vaccine.,
          State.of.Residence) %>% 
  mutate(State.of.Residence = recode_factor(.x = State.of.Residence,
                                        `International` = NA_character_)) %>% 
  filter(!is.na(State.of.Residence))

#Recode and label variables, ensure missing values are properly coded
covid.small <- covid.dirty %>% 
  select( Once.it.is.available..do.you.intend.to.get.the.vaccine.for.COVID.19.,
          Have.you.received.a.flu.shot.in.the.past.12.months., 
          What.is.the.highest.degree.or.level.of.school.you.have.completed., 
          Which.most.closely.aligns.with.your.political.beliefs.,
          I.am.at.risk.for.contracting.COVID.19.,
          COVID.19.can.be.a.serious.disease.,
          The.COVID.19.vaccine.will.prevent.me.from.contracting.COVID.19., 
          There.are.many.risks.associated.with.getting.the.COVID.19.vaccine.) %>% 
  mutate_all(as.factor) %>% 
  rename(vaccine.intention = Once.it.is.available..do.you.intend.to.get.the.vaccine.for.COVID.19.) %>% 
  rename(flu.shot.history = Have.you.received.a.flu.shot.in.the.past.12.months.) %>%
  mutate(flu.shot.history = recode_factor(.x = flu.shot.history,
                                          `Don't know/Not sure` = NA_character_)) %>% 
  rename(education.level = What.is.the.highest.degree.or.level.of.school.you.have.completed.) %>% 
  mutate(education.level = recode_factor(.x = education.level,
                                         '.' = NA_character_)) %>% 
  rename(political.beliefs = Which.most.closely.aligns.with.your.political.beliefs.) %>% 
  mutate(political.beliefs = recode_factor(.x = political.beliefs,
                                         '.' = "Prefer not to answer")) %>% 
  rename(perceived.suscept = I.am.at.risk.for.contracting.COVID.19.) %>% 
  mutate(perceived.suscept = recode_factor(.x = perceived.suscept,
                                         'Stongly Agree' = "Strongly Agree",
                                         'Stongly Disagree' = "Strongly Disagree",
                                         '.' = NA_character_)) %>% 
  rename(perceived.severity = COVID.19.can.be.a.serious.disease.) %>% 
  mutate(perceived.severity = recode_factor(.x = perceived.severity,
                                          'Stongly Agree' = "Strongly Agree",
                                          'Stongly Disagree' = "Strongly Disagree",
                                         `.` = NA_character_)) %>% 
  rename(perceived.benefit = The.COVID.19.vaccine.will.prevent.me.from.contracting.COVID.19.) %>%
  mutate(perceived.benefit = recode_factor(.x = perceived.benefit,
                                         `.` = NA_character_)) %>% 
  rename(perceived.barrier = There.are.many.risks.associated.with.getting.the.COVID.19.vaccine.) %>% 
  mutate(perceived.barrier = recode_factor(.x = perceived.barrier,
                                         `.` = NA_character_)) 

#summarize the cleaned data 
summary(covid.small)
```


```{r echo=FALSE, warning=FALSE, message=FALSE}

#Creating levels for each variable so they are ordered in logical sense on graphs below
covid.small$flu.shot.history <- factor(covid.small$flu.shot.history,
                                       levels = c("Yes", "No"))
covid.small$vaccine.intention <- factor(covid.small$vaccine.intention,
                                        levels = c("Yes", "No", "Don't Know/Not sure"))
covid.small$education.level <- factor(covid.small$education.level, 
                                      levels = c("Less than a high school diploma", "High school graduate, GED", "Some college/Associate's Degree", "Bachelor's Degree", "Graduate School"))
covid.small$political.beliefs <- factor(covid.small$political.beliefs, 
                                        levels = c("Very Conservative", "Somewhat Conservative", "Mixed", "Somewhat Liberal", "Very Liberal", "Prefer not to answer"))
covid.small$perceived.suscept <- factor(covid.small$perceived.suscept,
                                        levels = c("Strongly Agree", "Agree", "Neutral",
                                                  "Disagree", "Strongly Disagree"))
covid.small$perceived.severity <- factor(covid.small$perceived.severity,
                                        levels = c("Strongly Agree", "Agree", "Neutral",
                                                  "Disagree", "Strongly Disagree"))
covid.small$perceived.benefit <- factor(covid.small$perceived.benefit,
                                       levels = c("Strongly Agree", "Agree", "Neutral",
                                                  "Disagree", "Strongly Disagree"))
covid.small$perceived.barrier <- factor(covid.small$perceived.barrier,
                                        levels = c("Strongly agree", "Agree", "Neutral",
                                                  "Disagree", "Strongly disagree"))

```

# Descriptive Statistics
```{r echo=FALSE, warning=FALSE, message=FALSE}

#Open TableOne package 
library(package = "tableone")

#Creating table to show descriptive statistics in percents 
#add labels to the table
labelled::var_label(x = covid.small) <- c("Once it is available, do you intend to get the vaccine for COVID-19?",
                                          "Have you received a flu shot in the past 12 months?",
                                           "What is the highest degree or level of school you have completed?",
                                           "Which most closely aligns with your political beliefs?",
                                          "I am at risk for contracting COVID-19.",
                                          "COVID-19 can be a serious disease.",
                                          "The COVID-19 vaccine will prevent me from contracting COVID-19.",
                                          "There are many risks associated with getting the COVID-19 vaccine.")

# use print to show table with labels and percent
covid.table <- CreateTableOne(data = covid.small)
print(x = covid.table, 
      varLabels = TRUE,
      format = "p")

```


```{r echo=FALSE, message=FALSE, warning=FALSE}
#Figure 1 - Flu Shot History Graphed with Percent in Intention to get COVID-19 Vaccine
figure1 <- covid.small %>% 
  drop_na(vaccine.intention, flu.shot.history) %>% 
  group_by(vaccine.intention, flu.shot.history) %>% 
  count() %>% 
  group_by(flu.shot.history) %>% 
  mutate(percent = 100*(n/sum(n))) %>% 
  ggplot(aes(x = flu.shot.history, fill = vaccine.intention,
             y = percent)) +
  geom_col(position = "dodge") +
  coord_flip() + 
  theme_minimal() + 
  labs(x = "Have you received a flu shot in the past 12 months?", 
       y = "Percent in Flu Shot History",
       title = "Figure 1: Relationship between COVID Vaccine Intention \nand Flu Shot History") +
  scale_fill_manual(values = c("#f77d75", "#f5dfb8", "#76c1eb"), 
                    name = "COVID Vaccine Intention")
figure1
```


Flu Shot History Graphed with COVID-19 Vaccine Intention - Figure 1 shows the relationship between whether or not someone has gotten a flu shot in the past 12 months and their intention to get the COVID vaccine (n = 515). Of those who have gotten a flu vaccine in the past 12 months, approximately 58% of them intend to get the COVID-19 vaccine and 11% do not intend to get the COVID-19 vaccine. Of those who have not gotten a flu shot in the past 12 months, only 27% of them intend to get the COVID-19 vaccine and 35% do not. It can be inferred from this graph that those who have gotten a flu vaccine in the past 12 months are more likely to intend to get the COVID-19 vaccine. 


 
```{r echo=FALSE, message=FALSE, warning=FALSE}
#Figure 2 - Education Level Graphed with Percent in Intention to get COVID-19 Vaccine
figure2 <- covid.small %>% 
  drop_na(vaccine.intention, education.level) %>% 
  group_by(vaccine.intention, education.level) %>% 
  count() %>% 
  group_by(education.level) %>% 
  mutate(percent = 100*(n/sum(n))) %>%
  ggplot(aes(x = education.level, fill = vaccine.intention,
             y = percent)) +
  geom_col(position = "dodge") +
  coord_flip() +
  theme_minimal() + 
  labs(x = "Highest Achieved Education Level", 
       y = "Percent in Education Level",
         title = "Figure 2: Relationship between COVID Vaccine Intention \nand Highest Achieved Education Level") +
  scale_fill_manual(values = c("#f77d75", "#f5dfb8", "#76c1eb"),
                    name = "COVID Vaccine Intention")
figure2
```


Education Level Graphed with COVID-19 Vaccine Intention - Figure 2 shows the relationship between someone's highest achieved education level and their intention to get the COVID vaccine (n = 520). Of those who have achieved graduate school as their highest achieved educational level, approximately 57% intend to get the COVID-19 vaccine and 13% do not. Of those who have achieved high school as their highest achieved education level, approximately 20% intend to get the COVID-19 vaccine and 36% do not. This shows that higher achieved education levels may be associated with higher intention to get the COVID-19 vaccine.



```{r echo=FALSE, message=FALSE, warning=FALSE}
#Figure 3 - Political Beliefs Graphed with Percent in Intention to get COVID-19 Vaccine
figure3 <- covid.small %>% 
  drop_na(vaccine.intention, political.beliefs) %>% 
  group_by(vaccine.intention, political.beliefs) %>% 
  count() %>% 
  group_by(political.beliefs) %>% 
  mutate(percent = 100*(n/sum(n))) %>%
  ggplot(aes(x = political.beliefs, fill = vaccine.intention,
             y = percent)) +
  geom_col(position = "dodge") +
  coord_flip() +
  theme_minimal() + 
  labs(x = "Political Beliefs", 
       y = "Percent in Political Beliefs",
         title = "Figure 3: Relationship between COVID Vaccine Intention \nand Political Beliefs") +
  scale_fill_manual(values = c("#f77d75", "#f5dfb8", "#76c1eb"),
                    name = "COVID Vaccine Intention")
figure3
```


Political Beliefs Graphed with COVID-19 Vaccine Intention - Figure 3 shows the relationship between someone's political beliefs and their intention to get the COVID-19 vaccine (n = 521). The categories with the highest intention to get the COVID-19 vaccine are Very Liberal (74%) and Somewhat Liberal (52%). The categories with the highest amount of respondents who do not intend to get the COVID-19 vaccine are Prefer not to answer (65%) and Very Conservative (50%). This means that one's political beliefs likely has an impact on one's intention to get the COVID-19 vaccine. 



```{r echo=FALSE, message=FALSE, warning=FALSE}
#Figure 4 - Perceived Susceptibility Graphed with Percent in Intention to get COVID-19 Vaccine
figure4 <- covid.small %>% 
  drop_na(vaccine.intention, perceived.suscept) %>% 
  group_by(vaccine.intention, perceived.suscept) %>% 
  count() %>% 
  group_by(perceived.suscept) %>% 
  mutate(percent = 100*(n/sum(n))) %>%
  ggplot(aes(x = perceived.suscept, fill = vaccine.intention,
             y = percent)) +
  geom_col(position = "dodge") +
  coord_flip() + 
  theme_minimal() + 
  labs(x = "I am at risk for contracting COVID-19.", 
       y = "Percent in Perceived Susceptibility",
         title = "Figure 4: Relationship between COVID Vaccine Intention \nand One's Perceived Susceptibility") +
  scale_fill_manual(values = c("#f77d75", "#f5dfb8", "#76c1eb"),
                    name = "COVID Vaccine Intention")
figure4
```


Perceived Susceptibility Graphed with COVID-19 Vaccine Intention - Figure 4 shows the relationship between someone's perceived susceptibility (taken from the Health Belief Model) and their intention to get the COVID-19 vaccine (n = 520). Larger amounts of respondents Strongly Agree (57%) and Agree (59%) with the statement 'I am at risk for contracting COVID-19' and do intend to get the COVID-19 vaccine as compared to the other answer choices. Larger amounts of respondents Strongly Disagree (50%) and Disagree (33%) with the statement and do not intend to get the COVID-19 vaccine. This shows that respondents who believe they are susceptible to contracting COVID-19 may be more likely to intend to get the COVID-19 vaccine. 



```{r echo=FALSE, message=FALSE, warning=FALSE}
#Figure 5 - Perceived Severity Graphed with Percent in Intention to get COVID-19 Vaccine
figure5 <- covid.small %>% 
  drop_na(vaccine.intention, perceived.severity) %>% 
  group_by(vaccine.intention, perceived.severity) %>% 
  count() %>% 
  group_by(perceived.severity) %>% 
  mutate(percent = 100*(n/sum(n))) %>%
  ggplot(aes(x = perceived.severity, fill = vaccine.intention,
             y = percent)) +
  geom_col(position = "dodge") +
  coord_flip() + 
  theme_minimal() + 
  labs(x = "COVID-19 can be a serious disease.", 
       y = "Percent in Perceived Severity",
         title = "Figure 5: Relationship between COVID Vaccine Intention \nand One's Perceived Severity") +
  scale_fill_manual(values = c("#f77d75", "#f5dfb8", "#76c1eb"),
                    name = "COVID Vaccine Intention")
figure5
```


Perceived Severity Graphed with COVID-19 Vaccine Intention - Figure 5 shows the relationship between someone's perceived severity (taken from the Health Belief Model) and their intention to get the COVID-19 vaccine (n = 519). This graph is particularly interesting, as you can see a slight trend in the bars, but there are deviations away from that trend as well. According to the graph, more respondents who Strongly Agree (59%) with the statement 'COVID-19 can be a serious disease' do intend to get the COVID-19 vaccine; while a higher percentage of respondents who Disagree (88%) with the statement do not intend to get the COVID-19 vaccine. This shows that respondents who do not think COVID-19 is a serious disease are probably less likely to intend to get the COVID-19 vaccine. 



```{r echo=FALSE, message=FALSE, warning=FALSE}
#Figure 6 - Perceived Benefit Graphed with Percent in Intention to get COVID-19 Vaccine
figure6 <- covid.small %>% 
  drop_na(vaccine.intention, perceived.benefit) %>% 
  group_by(vaccine.intention, perceived.benefit) %>% 
  count() %>% 
  group_by(perceived.benefit) %>% 
  mutate(percent = 100*(n/sum(n))) %>%
  ggplot(aes(x = perceived.benefit, fill = vaccine.intention,
             y = percent)) +
  geom_col(position = "dodge") +
  coord_flip() + 
  theme_minimal() + 
  labs(x = "The COVID-19 vaccine will prevent \nme from contracting COVID-19.", 
       y = "Percent in Perceived Benefits",
         title = "Figure 6: Relationship between COVID Vaccine Intention \nand One's Perceived Benefits") +
  scale_fill_manual(values = c("#f77d75", "#f5dfb8", "#76c1eb"),
                    name = "COVID Vaccine Intention")
figure6
```


Perceived Benefit Graphed with COVID-19 Vaccine Intention - Figure 6 shows the relationship between someone's perceived benefits (taken from the Health Belief Model) and their intention to get the COVID-19 vaccine (n = 519). According to the graph, nearly 95% of respondents who Strongly Agree with the statement 'The COVID-19 vaccine will prevent me from contracting COVID-19' do intend to get the vaccine; while 88% of respondents who strongly disagree with that statement do not intend to get the vaccine. This means that it is likely that respondents who believe the vaccine works will be more likely to intend to get the COVID-19 vaccine. 



```{r echo=FALSE, message=FALSE, warning=FALSE}
#Figure 7 - Perceived Barrier Graphed with Percent in Intention to get COVID-19 Vaccine
figure7 <- covid.small %>% 
  drop_na(vaccine.intention, perceived.barrier) %>% 
  group_by(vaccine.intention, perceived.barrier) %>% 
  count() %>% 
  group_by(perceived.barrier) %>% 
  mutate(percent = 100*(n/sum(n))) %>%
  ggplot(aes(x = perceived.barrier, fill = vaccine.intention,
             y = percent)) +
  geom_col(position = "dodge") +
  coord_flip() + 
  theme_minimal() + 
  labs(x = "There are many risks associated with \ngetting the COVID-19 vaccine", 
       y = "Percent in Perceived Barriers",
       title = "Figure 7: Relationship between COVID Vaccine Intention \nand One's Perceived Barriers" ) +
  scale_fill_manual(values = c("#f77d75", "#f5dfb8", "#76c1eb"),
                    name = "COVID Vaccine Intention")
figure7
```


Perceived Barrier Graphed with COVID-19 Vaccine Intention - Figure 7 shows the relationship between someone's perceived barriers (taken from the Health Belief Model) and their intention to get the COVID-19 vaccine (n = 516). An overwhelming majority of respondents who Strongly Disagree (85%) and Disagree (85%) with the statement 'There are many risks associated with getting the COVID-19 vaccine' do intend to get the vaccine; while a majority of respondents who Strongly Agree with the statement (67%) do not intend to get the vaccine. This means it is probable that respondents will be less likely to get the vaccine if they believe there are risks associated with getting the vaccine. 




# Bivariate Tests and Logistic Regression Model 

### Logistic Regression
Logistic Regression is used to predict the probability of a categorical outcome variable, in this case, one's intention to get the COVID-19 vaccine. Because the outcome variable is not continuous, linear regression is not appropriate in this instance. For this research, we are determining whether one's flu shot history, political beliefs, highest achieved education, one's perceived severity, one's perceived susceptibility, one's perceived benefits, and one's perceived barriers help predict intention to get the COVID-19 vaccine. The outcome variable is a binary variable: yes, do intend to get the COVID-19 vaccine and no, do not intent to get the COVID-19 vaccine. Don't Know/Not Sure was also an option, but this was recoded as NA in order to create a binary outcome variable for regression modeling purposes. Because the outcome is binary, the model predicts probability of being in one of the two categories.

```{r include=FALSE , warning=FALSE, message=FALSE}
# Load libraries 
library(odds.n.ends)
library(car)

#Create another data set to create a binary outcome for regression purposes 
#Recode Less than a high school diploma as NA because it has zero responses and is messing with the P values 
#Collapse answer choices in order to meet the multicollinearity assumption
covid.binary <- covid.small %>% 
    mutate(vaccine.intention = recode_factor(.x = vaccine.intention,
                                          `Don't Know/Not sure` = NA_character_)) %>% 
    mutate(education.level = recode_factor(.x = education.level,
                                      `Less than a high school diploma` = NA_character_)) %>% 
    mutate(perceived.severity = recode_factor(.x = perceived.severity,
                                         'Strongly Agree' = "Agree",
                                         'Strongly Disagree' = "Disagree")) %>% 
    mutate(perceived.suscept = recode_factor(.x = perceived.suscept,
                                         'Strongly Agree' = "Agree",
                                         'Strongly Disagree' = "Disagree")) %>% 
   mutate(perceived.benefit = recode_factor(.x = perceived.benefit,
                                         'Strongly Agree' = "Agree",
                                         'Strongly Disagree' = "Disagree")) %>% 
   mutate(perceived.barrier = recode_factor(.x = perceived.barrier,
                                         'Strongly agree' = "Agree",
                                         'Strongly disagree' = "Disagree")) %>% 
  mutate(political.beliefs = recode_factor(.x = political.beliefs,
                                         'Very Conservative' = "Conservative",
                                         'Somewhat Conservative' = "Conservative",
                                         'Somewhat Liberal' = "Liberal",
                                         'Very Liberal' = "Liberal"))


#Summarize the cleaned data 
summary(covid.binary)

```


```{r include=FALSE , warning=FALSE, message=FALSE}
#relevel so that no is the reference group for the outcome variable
covid.binary <- covid.binary %>% 
  mutate(vaccine.intention = relevel(x = vaccine.intention, ref = "No"))
  
```

```{r include=FALSE , warning=FALSE, message=FALSE}
#check the level order (model predicts second level, first is reference)
levels(covid.binary$vaccine.intention)

```

```{r include=FALSE , warning=FALSE, message=FALSE}
#Check levels of predictors 
levels(covid.binary$flu.shot.history)
levels(covid.binary$education.level)
levels(covid.binary$political.beliefs)
levels(covid.binary$perceived.suscept)
levels(covid.binary$perceived.severity)
levels(covid.binary$perceived.benefit)
levels(covid.binary$perceived.barrier)

#Reference categories of each predictor
  # flu.shot.history - Yes 
  # education.level - High school graduate, GED
  # political.beliefs - Conservative
  # perceived.suscept - agree
  # perceived.severity - agree
  # perceived.benefit - agree
  # perceived.barrier - agree 

```

```{r echo=FALSE, warning=FALSE, message=FALSE}
#Bivariate tests (chi-squared) conducted prior to building a model

descTable <- CreateTableOne(data = covid.binary,
                            strata = 'vaccine.intention',
                            vars = c('flu.shot.history', 'education.level', 'political.beliefs', 'perceived.suscept', 'perceived.severity', 'perceived.benefit', 'perceived.barrier'))
print(descTable,
      varLabels = TRUE,
      format = "p")

```                           

There were statistically significant relationships between all predictors and the outcome, intention to get the COVID-19 vaccine (p < .001).


```{r echo=FALSE, warning=FALSE, message=FALSE}
#predict covid vaccine intention by the predictors 
covid.vaccine.model <- glm(vaccine.intention ~ flu.shot.history + education.level + political.beliefs + perceived.suscept + perceived.severity + 
                             perceived.barrier + perceived.benefit,
                           data = covid.binary,
                           family = "binomial")

odds.n.ends(covid.vaccine.model)

```

# Checking the Model Assumptions of Logistic Regression

### Assumption Check #1: Independence of Variables - NOT MET
This assumption is not met. Many of the respondents were friends or family of the author of the survey so this likely led to less variance in the data. 

### Assumption Check #2: Linearity - N/A
This assumption is not applicable to this model as their is no continuous variable included. 

### Assumption Check #3:  Multicollinearity - MET 
This assumption is met as all GVIF values are below 2. 

```{r echo=FALSE, warning=FALSE, message=FALSE}
# Multicollinearity test , values above two are not met 
vif(mod = covid.vaccine.model)

```

## Significant Odds Ratios: 
* People who identify as Somewhat Liberal and Very Liberal (combined to make the Liberal category) are 5.30 times more likely than people who identify as Somewhat Conservative and Very Conservative (combined to make the Conservative category) to intend to get the COVID-19 vaccine (OR = 5.30; 95% CI: 1.74 - 17.17).

* People who Disagree with the statement 'There are many risks associated with getting the COVID-19 vaccine' are 19.95 times more likely than people who agree with that statement to intend to get the COVID-19 vaccine (OR = 19.95; 95% CI:4.69 - 116.13).

* People who are Neutral on the statement 'There are many risks associated with getting the COVID-19 vaccine' are 3.53 times more likely than people who agree with that statement to intend to get the COVID-19 vaccine (OR = 3.53; 95% CI:1.38 - 9.46).

* The odds of someone intending to get the COVID-19 vaccine are 83% lower for people who have not gotten a flu shot in the past 12 months compared to people who have gotten a flu shot in the past 12 months (OR = .17; 95% CI: .07 - .42).

* The odds of someone intending to get the COVID-19 vaccine are 99% lower for people who disagree with the statement 'The COVID vaccine will prevent me from contracting COVID-19.' compared to people who agree with that statement (OR = .01; 95% CI: 0.00 - 0.06). 

* The odds of someone intending to get the COVID-19 vaccine are 96% lower for people who are neutral on the statement 'The COVID vaccine will prevent me from contracting COVID-19.' compared to people who agree with that statement (OR = .04; 95% CI: .01 - .14). 

* The odds of someone intending to get the COVID-19 vaccine are 95% lower for people who disagree with the statement 'I am at risk for contracting COVID-19.' compared to people who agree with that statement (OR = .05; 95% CI: .01 - .22). 


# Model Interpretation and Conclusion

A logistic regression model including flu shot history, educational level, political beliefs, perceived susceptibility, perceived severity, perceived benefits, and perceived barriers as predictors of one's intention to get the COVID-19 vaccine were statistically significantly better than the baseline at explaining the outcome [$\chi^2$(15) = 264.98; p < .05]. The model correctly predicted 227 of the 242 respondents who do intend to get the COVID-19 vaccine (93.8% correct - model sensitivity) and correctly predicted 83 of the 97 respondents who do not intend to get the COVID-19 vaccine (85.6% correct - model specificity) for a Count-R-squared of 0.898 or 89.8% correctly predicted. The model did meet the assumption of multicollinearity, but did not meet the assumption of independence of observations, so our results should be interpreted with caution and not generalized to the population. *Because assumption of multicollinearity failed initially, the researcher collapsed the answer choices for a few of the variables. All of the Likert scale questions (Perceived Susceptibility, Perceived Severity, Perceived Benefits, and Perceived Barriers) were collapsed from Strongly Agree, Agree, Neutral, Disagree, Strongly Disagree to Agree, Neutral, Disagree. Political Beliefs was also collapsed from Very Liberal, Somewhat Liberal, Mixed, Somewhat Conservative, and Very Conservative to Liberal, Mixed, and Conservative. After collapsing these variables, multicollinearity was met and the Odds Ratios were more digestible numbers.

The initial research question is 'What factors are associated with COVID-19 vaccine intention of adults in the United States?'. By just looking at the grouped bar graphs, all of the predictors seem to have some sort of relationship with the outcome, intention to get the COVID019 vaccine (n = 521). After conducting chi-square tests, it was found that there were statistically significant relationships between all predictors measured above and the outcome (p<.001). Using a logistic regression model to calculate the statistically significant Odds Ratios, people who identify as Liberal, people who do not think there are any risks associated with the COVID-19 vaccine, people who are neutral on if there are risks associated with the COVID-19 vaccine, people who have gotten a flu shot in the past 12 months, people who agree or are neutral on if the COVID-19 vaccine will prevent them from contracting COVID-19, and people who agree that they are at risk for contracting COVID-19 are all more statistically likely to intend to get the COVID-19 vaccine. (The exact odds for all of these groups are listed above.)

Although COVID-19 is a novel disease, my findings are very similar to most other vaccine intention studies. This information could help guide messaging campaigns regarding the COVID-19 vaccine at the local, state, and federal level. For example, based on my results, more messaging may need to be focused in counties that tend to lean conservative. Specific topics that should be focused on with vaccine messaging should be the benefits and barriers associated with getting vaccinated, as well as how susceptible one is for contracting COVID-19. Having this information can ensure messaging campaigns are well thought out and informative. For future research, the researcher believes it would be interesting to go into more detail on each health behavior construct in the survey. Furthermore, more questions could be added for each construct to have a better understanding of how the Health Belief Model constructs truly relate to COVID-19 vaccination intention. It would also be interesting to look at other factors from the survey in more detail, such as mask use, gender, state of residence, etc. These questions were asked on the survey, however, there was not a wide enough range of responses of these particular questions in order to do a true analysis on these specific variables. 


```{r echo=FALSE, warning=FALSE, message=FALSE}

#create codebook
library(dataMaid)
#makeCodebook(covid.small)

```

