---
title: "Two variable graphs"
author: "Jenine Harris"
date: "9/28/2021"
output: html_document
---

# FIX IT! Chapter 3 exercises 

This code is broken. Let's fix it.

```{r openPackages}
# load the tidyverse
library(package = "tidyverse")
library(package = "janitor")
```

The codebook for the 2011-2012 NHANES is on the CDC website: https://wwwn.cdc.gov/nchs/nhanes/continuousnhanes/default.aspx?BeginYear=2011

```{r importData}
# import the data (from local file, faster)
# uncomment next line to use this method to open
nhanes.2012 <- read.csv(file = "nhanes_2011_2012_ch3.csv")

# importing directly from the internet (more reproducible, slower)
# temp <- tempfile(fileext = ".zip")
# download.file(url = "https://edge.sagepub.com/system/files/datasets_7.zip", destfile = temp)
# nhanes.2012 <- read.csv(unz(temp, "Datasets/data/nhanes_2011_2012_ch3.csv")) %>% 
#   clean_names()
```

```{r cleanData}
# clean the income, marital status, race, and gun use variables
# choose legal adults with age > 17
nhanes.2012.clean <- nhanes.2012 %>% 
  select(indhhin2, dmdmartl, ridreth3, auq300, ridageyr, indfmpir) %>% 
  filter(ridageyr > 17) %>% 
  mutate(indhhin2 = recode_factor(indhhin2,
                                  `1` = "$0 to $4,999",
                                  `2` = "$5,000 to $9,999",
                                  `3` = "$10,000 to $14,999",
                                  `4` = "$15,000 to $19,999",
                                  `5` = "$20,000 to $24,999",
                                  `6` = "$25,000 to $34,999",
                                  `7` = "$35,000 to $44,999",
                                  `8` = "$45,000 to $54,999",
                                  `9` = "$55,000 to $64,999",
                                  `10` = "$65,000 to $74,999",
                                  `12` = "$20,000 and Over",
                                  `13` = "Under $20,000",
                                  `14` = "$75,000 to $99,999",
                                  `15` = "$100,000 and Over",
                                  `77` = NA_character_,
                                  `99` = NA_character_)) %>% 
  mutate(dmdmartl = recode_factor(dmdmartl,
                                  `1` = 'Married',
                                  `2` = 'Widowed',
                                  `3` = 'Divorced',
                                  `4` = 'Separated',
                                  `5` = 'Never married',
                                  `6` = 'Living with partner',
                                  `77` = NA_character_,
                                  `99` = NA_character_)) %>% 
  mutate(ridreth3 = recode_factor(ridreth3, 
                                  `1` = "Mexican American",
                                  `2` = "Other Hispanic", 
                                  `3` = "Non-Hispanic White",
                                  `4` = "Non-Hispanic Black",
                                  `6` = "Non-Hispanic Asian",
                                  `7` = "Other Race - Including Multi-Racial")) %>% 
  mutate(auq300 = recode_factor(auq300, 
                                `1` = "Yes",
                                `2` = "No",
                                `7` = NA_character_)) %>% 
  rename(race.eth = ridreth3) %>%  
  rename(income = indhhin2) %>%  
  rename(marital = dmdmartl) %>% 
  rename(gun.use = auq300) %>% 
  rename(age = ridageyr) %>% 
  rename(poverty.ratio = indfmpir) %>% 
  drop_na()

summary(object = nhanes.2012.clean)
table(nhanes.2012.clean$income)
```

```{r incomeGraph}
# Income graph
nhanes.2012.clean %>%
  #drop_na() %>% 
  ggplot(aes(x = income)) +
  geom_bar() +
  labs(x = "Income", y = "Number of observations",
       title = "Income category of adult survey participants", 
       subtitle = "NHANES 2011-2012 (n = 4,213)") 
```

Yikes! Make a formatting adjustment to be able to read the categories.

3 things that should be in every table/figure title:
- sample size 
- year of data collection
- source of data (this could also go in a footnote)

What is going on with the income variable?

What could we do to fix it?


## Bivariate graphs

### Two categorical

```{r incomeGunGraph}
# Income & gun use graph (with number of observations)
nhanes.2012.clean %>%
  ggplot(aes(x = income, fill = gun.use)) +
  geom_bar(position = "dodge") + # position = "dodge" puts bars side-by-side
  coord_flip() +
  labs(x = "Income category", y = "Number of observations",
       fill = "Ever used gun",
       title = "Adults ever used gun by income (n = 4,213)", 
       subtitle = "2011-2012 NHANES") +
  theme_minimal() +
  scale_fill_brewer(palette = "Reds")
```

Interpret the graph: 

```{r incomeGunPercentGraph}
# Income & gun use graph (with percentages)
nhanes.2012.clean %>%
  drop_na(gun.use, income) %>% 
  ggplot(aes(x = income, fill = gun.use,
             y = 100*(..count..)/sum(..count..))) +
  geom_bar(position = "dodge") + # position = "dodge" puts bars side-by-side
  coord_flip() +
  labs(x = "Income category", y = "Percent of observations",
       fill = "Gun use",
       title = "Gun use by income for adult participants\nof 2011-2012 NHANES (n = 4,213)") +
  theme_minimal() +
  scale_fill_brewer(palette = "Reds")
```

How are these percentages being computed? 



Which bars add up to 100%?




```{r incomeGunPercentCorrected}
# Income & gun use graph (with percentages)
nhanes.2012.clean %>% 
  group_by(income, gun.use) %>% 
  count() %>% 
  group_by(income) %>% # the variable you want to add up to 100%
  mutate(percentInc = 100*(n/sum(n))) %>% 
  ggplot(aes(x = income, fill = gun.use,
             y = percentInc)) +
  geom_col(position = "dodge") +
  coord_flip() +
  labs(x = "Income category", y = "Percent within income group", 
       fill = "Ever used gun") +
  theme_minimal()+
  scale_fill_brewer(palette = "Reds")
```

How are these percentages being computed? 


Which bars add up to 100%?



```{r raceGun}
# Race-ethnicity & gun use graph with percentages
nhanes.2012.clean %>%
  drop_na(race.eth, gun.use) %>% 
  group_by(race.eth, gun.use) %>% 
  count() %>% 
  group_by(race.eth) %>% # the variable you want to add to 100%
  mutate(perc.gun.use = 100*(n/sum(n))) %>% 
  ggplot(aes(x = race.eth, y = perc.gun.use, fill = gun.use)) +
  geom_col(position = "dodge") +
  coord_flip() +
  labs(x = "Race-ethnicity", y = "Number of observations") +
  theme_minimal() +
  scale_fill_brewer(palette = "Paired", name = "Ever used gun")
```

Interpret the race/ethnicity graph: 

### One categorical and one continuous

```{r ageGun}
# Age & gun use graph
nhanes.2012.clean %>%
  ggplot(aes(x = age, y = gun.use)) +
  geom_boxplot(fill = "deeppink") +
  labs(x = "Age", y = "Ever gun use") +
  theme_minimal() 
```

```{r ageGunHist}
# Age & gun use graph 
nhanes.2012.clean %>%
  ggplot(aes(x = age)) +
  geom_histogram(bins = 10, fill = "deeppink", color = "white") +
  facet_wrap(facets = vars(gun.use)) +
  labs(x = "Age", y = "Frequency") +
  theme_minimal() 
```

Interpret the age graphs:


### Two continuous

*Note: The poverty ratio measures the ratio of family income to poverty. When the ratio is 1, the income and poverty level are the same. When the ratio is below 1, income is below poverty level. Ratios above 1 are above poverty level. Generally, the higher the poverty ratio, the more the income.

```{r agePoverty}
# Age & poverty ratio graph 
# choose under 25 age
nhanes.2012.clean %>%
  filter(age < 25) %>% 
  ggplot(aes(x = age, y = poverty.ratio)) +
  geom_jitter(color = "gold", size = 3) +
  labs(x = "Age", y = "Poverty ratio") +
  theme_minimal() 
```

Interpret the age and poverty graph:


## You try it!

```{r maritalGun}
# Marital status & gun use graph with percentages

```

Interpret the marital status graph: 


## The End