---
title: "Week 3 Workshop: Data Cleaning and Making Tables"
subtitle: "Are sex, gender identity, and age associated with heavy drinking? A mystery with broken code"
output: html_document
---

# Instructions 

* Wherever you see a ________, replace it with whatever is needed for the code to work! Find these blanks by looking for the red x next to the line number on the left. The red x indicates something is wrong with the code and is a great way to debug code that is not running as expected.

* Wherever you see a question number in parentheses like this (3), answer the question by typing your answer directly into this document.

* Remember to save your work often!

# Opening the R packages needed for this workshop

```{r openPackages}
# if needed, install packages through the Tools menu: tidyverse, haven, table1
# load packages needed for this workshop
library(________ = "tidyverse")
library(package = ________)
________(package = "table1")
```

# Importing the data

* The data can be found on the CDC website and downloaded directly into R from there

Behavioral Risk Factor Surveillance Survey (BRFSS)

* The file is an xpt file (SAS transport file type) so the import function is `read_xpt`

```{r importData}
# 2015 Behavioral Risk Factor Surveillance Survey data
# opening directly from the online location
# this may take a few minutes depending on your computer and internet speed
temp <- tempfile(fileext = ".zip")
download.file(url  = "https://www.cdc.gov/brfss/annual_data/2015/files/LLCP2015XPT.zip",
              destfile = temp)
BRFSS_2015 <- read_xpt(file = temp)
```

# Understanding the variables of interest

* We will answer the research question: 
    + Are sex, gender identity, and age associated with heavy drinking? 
    
* We will use the following variables:
    + SEX
    + TRNSGNDR
    + _RFDRHV5
    + _DRNKWEK
    + _AGE80
    
* Search the codebook to learn how these variables were measured:
    + https://www.cdc.gov/brfss/annual_data/2015/pdf/codebook15_llcp.pdf
    
(1) Describe how the variables were measured:

  (a) SEX: 
  
  (b) TRNSGNDR: 
  
  (c) _RFDRHV5: 
  
  (d) _DRNKWEK: 
  
  (e) _AGE80: 

(2) What are the descriptive statistics that would be appropriate for each variable:

  (a) SEX: 
  
  (b) TRNSGNDR: 
  
  (c) _RFDRHV5: 
  
  (d) _DRNKWEK: 
  
  (e) _AGE80: 


# Select the variables of interest

```{r selectVariable}
# select variables
# back ticks around names of variables that start with a special character
BRFSS_2015_small <- BRFSS_2015 %>%
   select(SEX, ________, `_RFDRHV5`, `_DRNKWEK`, ________)
```

# Examine the variables of interest

```{r variableSummary}
# summarize small data set
________(object = BRFSS_2015_small)
```

(3) Based on this, what is a good data management plan:



# Recoding the sex variable 

```{r recodeSex}
# add on to the previous data management
BRFSS_2015_small <- BRFSS_2015 %>%
  select(SEX, TRNSGNDR, `_RFDRHV5`, `_DRNKWEK`, `_AGE80`) %>% 
  mutate(SEX = recode_factor(________,
                             '2' = ________,
                             '1' = 'male'))
  
```

# Checking the work so far

```{r summarizeData}
# summarize the data frame
summary(object = ________)
```

# Recoding the transgender variable

```{r recodeTrans}
# add on to the previous data management
BRFSS_2015_small <- BRFSS_2015 %>%
  select(SEX, TRNSGNDR, `_RFDRHV5`, `_DRNKWEK`, `_AGE80`) %>%  
  mutate(SEX = recode_factor(SEX,
                             '2' = 'female',
                             '1' = 'male')) %>% 
  mutate(TRNSGNDR = recode_factor(________,
                                  '1' = 'male-to-female',
                                  '2' = 'female-to-male',
                                  '3' = ________,
                                  '4' = 'not transgender',
                                  '7' = NA_character_,
                                  '9' = ________)) 
  
```

# Checking the work so far

```{r summarizeData2}
# summarize the data frame
summary(object = BRFSS_2015_small)
```

# You try it!

* Fill in the code to recode the heavy drinking variable `_RFDRHV5` 

```{r recodeHeavyDrink}
# add on to the previous data management
BRFSS_2015_small <- BRFSS_2015 %>%
  select(SEX, TRNSGNDR, `_RFDRHV5`, `_DRNKWEK`, `_AGE80`) %>%  
  mutate(SEX = recode_factor(SEX,
                             '2' = 'female',
                             '1' = 'male')) %>% 
  mutate(TRNSGNDR = recode_factor(TRNSGNDR,
                                  '1' = 'male-to-female',
                                  '2' = 'female-to-male',
                                  '3' = 'gender non-conforming',
                                  '4' = 'not transgender',
                                  '7' = NA_character_,
                                  '9' = NA_character_)) %>% 
  mutate(________ = recode_factor(________,
                                    '1' = ________,
                                    '2' = ________,
                                    '9' = ________))
```

# Checking the work so far

```{r summarizeData3}
# summarize the data frame
________(object = ________)
```

# Recoding `_DRNKWEK`, renaming stuff, dropping observations with missing values 

```{r recodeDrnwek}
# add on to the previous data management
BRFSS_2015_small <- BRFSS_2015 %>%
  select(SEX, TRNSGNDR, `_RFDRHV5`, `_DRNKWEK`, `_AGE80`) %>%  
  mutate(SEX = recode_factor(________,
                             '2' = 'female',
                             '1' = 'male')) %>% 
  mutate(TRNSGNDR = recode_factor(________,
                                  '1' = 'male-to-female',
                                  '2' = 'female-to-male',
                                  '3' = 'gender non-conforming',
                                  '4' = 'not transgender',
                                  '7' = NA_character_,
                                  '9' = NA_character_)) %>% 
  mutate(`_RFDRHV5` = recode_factor(________,
                                    '1' = 'no',
                                    '2' = 'yes',
                                    '9' = NA_character_)) %>% 
  mutate(`_DRNKWEK` = `_DRNKWEK`/100) %>% 
  mutate(`_DRNKWEK` = na_if(`_DRNKWEK`, 999)) %>% 
  rename(sex = SEX) %>% 
  rename(genderIdentity = ________) %>% 
  rename(heavyDrink = `_RFDRHV5`) %>% 
  rename(drinksWeek = ________) %>% 
  rename(ageYears = `_AGE80`) %>% 
  drop_na()
# drop_na() with empty parentheses 
# drops any observation with an NA value for any variable
```

# Checking the work so far

```{r summarizeData4}
# summarize the data set
summary(object = BRFSS_2015_small)
```

# Choosing appropriate descriptive statistics

```{r histograms}
# check the continuous variable distributions
# plot the variable distribution to choose appropriate descriptive stats
BRFSS_2015_small %>% 
  ggplot(aes(x = drinksWeek)) +
  geom_histogram()
# Not normal, right skewed, use median/IQR

________ %>% 
  ________(aes(x = ageYears)) +
  ________()
# skewed left, use median/IQR
```
(4) Describe the distribution and which descriptive statistics would be good for these two variables:

  (a) drinksWeek: 
  
  (b) ageYears: 

# Make a table of descriptive statistics

```{r table}
# table of all variables
table1(~ sex + genderIdentity + heavyDrink + ________ + ________,
       data = BRFSS_2015_small)

```

# Add labels to the table

```{r labels}
# labels for each variable
label(BRFSS_2015_small$sex) = "Participant sex"
label(_____________$genderIdentity) = "Participant gender identity"
label(BRFSS_2015_small$heavyDrink) = "Participant is heavy drinker"
label(_____________$________) = "Number of alcohol drinks per week"
label(_____________$________) = "Participant age in years"
```

# Print the table with labels

```{r tableWithLabels}
# table of all variables with labels
table1(~ sex + genderIdentity + heavyDrink + drinksWeek + ageYears,
       data = BRFSS_2015_small)
```

# Specify the appropriate descriptive statistics in the table

* Change the statistics displayed for the continuous variables

```{r correctTableWithLabels}
# table of all variables with labels
table1(~ sex + genderIdentity + heavyDrink + drinksWeek + ageYears, 
       render.continuous = c(. = "median (IQR)"),
       ________ = ________)
```


# Answer the research question

```{r stratifiedTable}
# table of gender identity and age by heavy drinker and sex
table1(~ genderIdentity + ________ | sex * heavyDrink, 
       render.continuous = c(. = "median (IQR)"),
       ________ = BRFSS_2015_small)
```

(5) Interpret the descriptive statistics for between heavy drinking and the three variables of interest:

  (a) Heavy drinking & sex: 
  
  (b) Heavy drinking & gender identity: 
  
  (c) Heavy drinking & age: 


# EXTRA STUFF (ON YOUR OWN): Examining the relationships using graphs

## Heavy drinking by sex

```{r barDrinkingBySex}
# graphing frequencies 
BRFSS_2015_small %>% 
  ggplot(aes(x = ________, fill = sex)) +
  geom_bar(position = "dodge") +
  theme_minimal() +
  labs(x = "Participant is heavy drinker", 
       fill = "Participant sex") 
```

```{r percentBarDrinkSex}
# graphing percentages (see page 48 in book & chapter 3) 
BRFSS_2015_small %>% 
  group_by(heavyDrink, sex) %>% 
  count() %>% 
  group_by(heavyDrink) %>% 
  mutate(perc.drink = 100*n/sum(n)) %>% 
  ggplot(aes(x = heavyDrink, 
             y = perc.drink,
             fill = sex)) +
  geom_col(position = "dodge") +
  theme_minimal() +
  labs(x = ________, 
       y = ________, 
       fill = ________) 
```

## Gender identity by heavy drinking

```{r identityDrinking}
# graphing the frequencies
BRFSS_2015_small %>% 
  ggplot(aes(x = heavyDrink, 
             fill = genderIdentity)) +
  geom_bar(position = "dodge") +
  theme_minimal() +
  labs(x = ________, 
       y = ________, 
       fill = ________) 
```



## Age by heavy drinking

```{r ageDrinking}
BRFSS_2015_small %>% 
  ggplot(aes(x = ageYears, 
             fill = heavyDrink)) +
  geom_histogram(color = 'white') +
  theme_minimal() +
  labs(x = ________, 
       y = ________, 
       fill = ________) 
```



# BONUS MYSTERY (EXPLORE ON YOUR OWN): What is the sex variable actually measuring, though?

* Is the sex variable measuring sex recorded at birth? or
* Is the sex variable measuring gender identity? 

```{r genderSex}
# make a table of gender identity by sex
table1(~ genderIdentity | sex, 
       data = BRFSS_2015_small)
```

(6) Interpret the table: 


# Make a visual to aid in interpretation

```{r barGenderSex}
# create a bar plot showing sex by gender identity
BRFSS_2015_small %>% 
  ggplot(aes(x = sex, 
             fill = ________)) +
  geom_bar()

```

# Filter out the "not transgender" group to see more clearly

```{r barGenderSexOnlyTrans}
# create a bar plot showing sex by gender identity
BRFSS_2015_small %>% 
  filter(genderIdentity != "not transgender") %>% 
  ggplot(aes(x = sex, 
             fill = ________)) +
  geom_bar()

```

(7) Interpret this graph:


# Why does this even matter for our research question?

* Check the codebook to see how the heavy drinking variable was created

(8) Complete this statement: Heavy drinking was defined as...


# What are some options for next steps? 

* Ideally, learn more about how recorded sex at birth and current sex or gender identity are associated with the health outcomes from heavy drinking and use this information to guide data management and analyses 

* Operationally define the sex variable as sex recorded at birth or as current sex 

* If choosing to define sex as sex recorded at birth, recode:
    + Participants who chose male-to-female gender identity and female sex to male
    + Participants who chose female-to-male gender identity and male sex to female
    
* If choosing to define sex as current sex or gender identity, recode:
    + Participants who chose male-to-female gender identity and male sex to female 
    + Participatns who chose female-to-male gender identity and female sex to male 

* What about gender non-conforming? Leave in as-is? Remove? 


# Recoding the sex variable (Beware! Complex code!)

The data management plan:
* Everyone with "not transgender" gender identity stays with the current value of sex
* Depending on the choice above, people with transgender gender identities may be recoded for sex recorded at birth or current sex 

EDIT THE CODE TO KEEP EITHER THE sexAtBirth VARIABLE -OR- THE sexCurrent VARIABLE and delete the other

```{r chooseCode}
# recoding sex variable 
# dropping the gender non-conforming with the filter() function
# removing the gender non-conforming category with the droplevels() function
BRFSS_2015_small_sex_recoded <- BRFSS_2015_small %>% 
  filter(genderIdentity != "gender non-conforming") %>% 
  mutate(sexAtBirth = case_when(genderIdentity == "female-to-male" ~ "female",
                                genderIdentity == "male-to-female" ~ "male",
                                sex == "female" & genderIdentity == "not transgender" ~ "female",
                                sex == "male" & genderIdentity == "not transgender" ~ "male")) %>%
  mutate(sexCurrent = case_when(genderIdentity == "female-to-male" ~ "male",
                                genderIdentity == "male-to-female" ~ "female",
                                sex == "female" & genderIdentity == "not transgender" ~ "female",
                                sex == "male" & genderIdentity == "not transgender" ~ "male")) %>%
  droplevels()

```

# Check the recoding 

EDIT THE CODE TO KEEP EITHER THE TABLE FOR THE sexAtBirth VARIABLE -OR- THE sexCurrent VARIABLE and delete the other

```{r tables}
# check your work with a table of gender identity by sex
table1(~ genderIdentity | sexAtBirth, 
       data = BRFSS_2015_small_sex_recoded)

# check your work with a table of gender identity by sex
table1(~ genderIdentity | sexCurrent, 
       data = BRFSS_2015_small_sex_recoded)
```

(9) Interpret the statistics in the table:

# Graph something!

* Pick something you are interested in about the work we did today and make a graph about it
* Some ideas:
    + Make some basic bar graphs for each of the variables in the data frame
    + Graph the recoded sex variable vs. the gender identity variable to check the work 
    + Graph the recoded sex variable vs. the heavy drinker variable to see any differences from the previous graph
    + Graph the recoded sex variable vs. the heavy drinker variable and add facets for the original sex variable (add this facet layer: facet_grid(cols = vars(sex)) ...see https://ggplot2.tidyverse.org/reference/facet_grid.html for more ideas)

```{r graphSomething}

```


(10) Interpret your graph(s):


# The end!


