---
title: "Chap 3 exercises"
author: "Jenine Harris"
date: "9/28/2021"
output: html_document
---

# FIX IT! Chapter 3 exercises 

This code is broken. Let's fix it.

```{r openPackages}
# load the tidyverse
library(package = "tidyverse")
```

The codebook for the 2011-2012 NHANES is on the CDC website: https://wwwn.cdc.gov/nchs/nhanes/continuousnhanes/default.aspx?BeginYear=2011

```{r importData}
# 1) Income

# import the data (from local file, faster)
# uncomment next line to use this method to open
# nhanes.2012 <- read.csv(file = "nhanes_2011_2012_ch3.csv")

# importing directly from the internet (more reproducible, slower)
temp <- tempfile(fileext = ".zip")
download.file(url = "https://edge.sagepub.com/system/files/datasets_7.zip", destfile = temp)
nhanes.2012 <- read.csv(unz(temp, "Datasets/data/nhanes_2011_2012_ch3.csv"))
```

```{r cleanData}
# clean the income, marital status, race, and gun use variables
nhanes.2012.clean <- nhanes.2012 %>% 
  select(INDHHIN2, DMDMARTL, RIDRETH3, AUQ300) %>% 
  mutate(INDHHIN2 = recode_factor(INDHHIN2,
                                  `1` = "$0 to $4,999",
                                  `2` = "$5,000 to $9,999",
                                  `3` = "$10,000 to $14,999",
                                  `4` = "$15,000 to $19,999",
                                  `5` = "$20,000 to $24,999",
                                  `6` = "$25,000 to $34,999",
                                  `7` = "$35,000 to $44,999",
                                  `8` = "$45,000 to $54,999",
                                  `9` = "$55,000 to $64,999",
                                  `10` = "$65,000 to $74,999",
                                  `12` = "$20,000 and Over",
                                  `13` = "Under $20,000",
                                  `14` = "$75,000 to $99,999",
                                  `15` = "$100,000 and Over",
                                  `77` = NA_character_,
                                  `99` = NA_character_)) %>% 
  mutate(DMDMARTL = recode_factor(DMDMARTL,
                                  `1` = 'Married',
                                  `2` = 'Widowed',
                                  `3` = 'Divorced',
                                  `4` = 'Separated',
                                  `5` = 'Never married',
                                  `6` = 'Living with partner',
                                  `77` = NA_character_,
                                  `99` = NA_character_)) %>% 
  mutate(RIDRETH3 = recode_factor(RIDRETH3, 
                                  `1` = "Mexican American",
                                  `2` = "Other Hispanic", 
                                  `3` = "Non-Hispanic White",
                                  `4` = "Non-Hispanic Black",
                                  `6` = "Non-Hispanic Asian",
                                  `7` = "Other Race - Including Multi-Racial")) %>% 
  mutate(AUQ300 = recode_factor(AUQ300, 
                                `1` = "Yes",
                                `2` = "No",
                                `7` = NA_character_)) %>% 
  rename(race.eth = RIDRETH3) %>%  
  rename(income = INDHHIN2) %>%  
  rename(marital = DMDMARTL) %>% 
  rename(gun.use = AUQ300) %>% 
  drop_na()

summary(object = nhanes.2012.clean)
table(nhanes.2012.clean$income)
```

```{r incomeGraph}
# Income graph
nhanes.2012.clean %>%
  #drop_na() %>% 
  ggplot(aes(x = income)) +
  geom_bar() +
  coord_flip() +
  labs(x = "Income category", y = "Number of observations",
       title = "Income distribution in 2011-2012 NHANES data (n = 4,213).") 
```

3 things that should be in every table/figure title:
- sample size 
- year of data collection
- source of data (this could also go in a footnote)

What is going on with the income variable?

What could we do to fix it?

```{r maritalGraph}
# 2) Marital status
nhanes.2012.clean %>% 
  #drop_na() %>% 
  ggplot(aes(x = marital)) + 
  geom_bar() +
  labs(x = "Marital status", y = "Number of observations",
       title = "Marital status of survey participants", 
       subtitle = "NHANES 2011-2012 (n = 4,213)") +
  coord_flip() +
  theme_bw()
```

Interpret the graph: 

```{r raceGraph}
# 3) Race
nhanes.2012.clean %>%   
  #drop_na() %>% 
  ggplot(aes(x = race.eth)) + 
  geom_bar() +
  coord_flip() +
  labs(x = "Race / ethnicity", y = "Number of observations",
       title = "Race / ethnicity of survey participants",
       subtitle = "NHANES 2011-2012 (n = 4,213)")

nhanes.2012.clean %>%   
  #drop_na() %>% 
  ggplot(aes(x = race.eth)) + 
  geom_bar() +
  coord_flip() +
  labs(x = "Race and Ethinicty", y ="Number of observations",
       title = "Race and Ethnicity of survey participants",
       subtitle = "NHANES 2011-2012 (n = 4,213)")+
  coord_flip() +
  theme_minimal()

```

Interpret the graph: 

```{r incomeGunGraph}
# Income & gun use graph (with number of observations)
nhanes.2012.clean %>%
  ggplot(aes(x = income, fill = gun.use)) +
  geom_bar(position = "dodge") + # position = "dodge" puts bars side-by-side
  coord_flip() +
  labs(x = "Income category", y = "Number of observations",
       fill = "Gun use",
       title = "Gun use by income for participants\nof 2011-2012 NHANES (n = 4,213)") +
  theme_minimal() +
  scale_fill_brewer(palette = "Reds")
```

Interpret the graph: 

```{r incomeGunPercentGraph}
# Income & gun use graph (with percentages)
nhanes.2012.clean %>%
  drop_na(gun.use, income) %>% 
  ggplot(aes(x = income, fill = gun.use,
             y = 100*(..count..)/sum(..count..))) +
  geom_bar(position = "dodge") + # position = "dodge" puts bars side-by-side
  coord_flip() +
  labs(x = "Income category", y = "Percent of observations",
       fill = "Gun use",
       title = "Gun use by income for participants\nof 2011-2012 NHANES (n = 4,213)") +
  theme_minimal() +
  scale_fill_brewer(palette = "Reds")
```

What sort of percentages are these? 



Which bars add up to 100%?




```{r incomeGunPercentCorrected}
# Income & gun use graph (with percentages)
nhanes.2012.clean %>% 
  group_by(income, gun.use) %>% 
  count() %>% 
  group_by(income) %>% # the variable you want to add up to 100%
  mutate(percentInc = 100*(n/sum(n))) %>% 
  ggplot(aes(x = income, fill = gun.use,
             y = percentInc)) +
  geom_col(position = "dodge") +
  coord_flip() +
  labs(x = "Income category", y = "Percent within income group", 
       fill = "Ever used gun") +
  theme_minimal()+
  scale_fill_brewer(palette = "Reds")
```

What sort of percentages are these? 

Which bars add up to 100%?



```{r raceGun}
# Race-ethnicity & gun use graph with percentages
nhanes.2012.clean %>%
  drop_na(race.eth, gun.use) %>% 
  group_by(race.eth, gun.use) %>% 
  count() %>% 
  group_by(race.eth) %>% 
  mutate(perc.gun.use = 100*(n/sum(n))) %>% 
  ggplot(aes(x = race.eth, y = perc.gun.use, fill = gun.use)) +
  geom_col(position = "dodge") +
  coord_flip() +
  labs(x = "Race-ethnicity", y = "Number of observations") +
  theme_minimal() +
  scale_fill_brewer(palette = "Paired", name = "Ever used gun")
```



Interpret the graph: 


```{r maritalGun}
# Marital status & gun use graph with percentages

```

Interpret the graph: 


## The End