---
title: 'Explaining or predicting zombies'
output: html_document
---

## Purpose

Use chi-squared and t-tests to determine which characteristics are associated with being a zombie. 

Practice the chi-squared and t-test process and interpretation.

```{r loadPackagesData, include=FALSE}
# import the zombie data
# open needed packages 
# make sure you have tidyverse, table1, haven, descr, and labelled packages
library(package = "tidyverse")
library(package = "table1")
library(package = "labelled")
library(package = "haven")
library(package = "descr")
# import data
zombies <- read_dta("zombies.dta")
```

## We need a codebook

```{r makeCodebook, include=FALSE}
codebook.zombies <- labelled::generate_dictionary(data = zombies)
codebook.zombies
```

## Clean data

```{r cleanData}
# cleaning the data 
zombies <- zombies %>% 
  mutate(zombie = recode_factor(.x = zombie,
                                '0' = "Not zombie",
                                '1' = "Zombie")) %>% 
  mutate(waterPerPerson = water / household) %>% 
  mutate_if(is.character, as.factor)
```

## Chi-squared 

Chi-squared is used to determine whether there is a relationship between two categorical variables (PAGE: 277, 291)

In determining the characteristics associated with being a zombie, which is a categorical variable, which of the variables below would be examined using chi-squared (delete the ones that are not):

* **gender:** sex (female/male)
* **rurality:** lives in rural/suburban/urban area
* **radio:** have a battery powered radio (yes/no)
* **flashlight:** have a flashlight (yes/no)
* **firstaid:** have a first aid kit (yes/no)

## What are the assumptions for chi-squared? (PAGE: 293)

* Variables may be nominal or ordinal
* Expected values should be 5 or higher in at least 80% of groups
* Observations must be independent

## Conducting chi-squared

## Make a table to explore the data first

```{r table}
table1(~ gender + rurality + radio + flashlight + firstaid + age | zombie,
       render.continuous = "Median (IQR)",
       data = zombies)
```

Based just on these descriptive statistics, were does it look like there may be a relationship between zombie status and some characteristic? 


### Sex and zombie status

H0: There is no relationship between sex and zombie status

HA: There is a relationship between sex and zombie status

#### Make a prediction about what you will find based on data visualization 

```{r sexZombieBarGraph}
# sex & zombie status (see page 289)
zombies %>% 
  group_by(gender, zombie) %>% 
  count() %>% 
  group_by(gender) %>% 
  mutate(perc.gender = 100*n/sum(n)) %>% 
  ggplot(aes(x = gender, y = perc.gender, fill = zombie)) +
  geom_col(position = "dodge") +
  theme_minimal() +
  labs(y = "Percent of zombies within gender group",
       x = "Gender")
```

Prediction: 

#### Conduct the analysis

```{r sexZombieChiSq}
# sex and zombie status
CrossTable(y = zombies$zombie,
           x = zombies$gender,
           expected = TRUE,
           prop.c = FALSE,
           prop.t = FALSE,
           prop.chisq = FALSE,
           chisq = TRUE,
           sresid = TRUE)
```

Does it meet assumptions and why or why not?: 

.

Interpret the results: 



### Rurality and zombie status

H0: There is no relationship between rurality and zombie status.

HA: There is a relationship between rurality and zombie status.

#### Make a prediction about what you will find based on exploratory analysis: 

```{r ruralZombieGraph}
# rural & zombie status (see page 289)
zombies %>% 
  group_by(rurality, zombie) %>% 
  count() %>% 
  group_by(rurality) %>% 
  mutate(perc.rurality = 100*n/sum(n)) %>% 
  ggplot(aes(x = rurality, y = perc.rurality, fill = zombie)) +
  geom_col(position = "dodge") +
  theme_minimal() +
  labs(y = "Percent of zombies within rurality group",
       x = "Rurality")
```

Prediction: 

#### Conduct the analysis

```{r ruralZombieChiSq}
#  rurality and zombie status
CrossTable(y = zombies$zombie,
           x = zombies$rurality,
           expected = TRUE,
           prop.c = FALSE,
           prop.t = FALSE,
           prop.chisq = FALSE,
           chisq = TRUE,
           sresid = TRUE)
```

Does it meet assumptions and why or why not?: 



Interpret the results:



### Radio and zombie status

H0: There is no relationship between having a radio and zombie status

HA: There is a relationship between having a radio and zombie status

#### Make a prediction about what you will find based on exploratory analysis: 

```{r radioZombieBarGraph}
# radio & zombie status (see page 289)
zombies %>% 
  group_by(radio, zombie) %>% 
  count() %>% 
  group_by(radio) %>% 
  mutate(perc.radio = 100*n/sum(n)) %>% 
  ggplot(aes(x = radio, y = perc.radio, fill = zombie)) +
  geom_col(position = "dodge") +
  theme_minimal() +
  labs(y = "Percent of zombies within radio group",
       x = "Has radio")
```

Prediction: 

#### Conduct the analysis

```{r radioZombieChiSq}
# radio and zombie status
CrossTable(y = zombies$zombie,
           x = zombies$radio,
           expected = TRUE,
           prop.c = FALSE,
           prop.t = FALSE,
           prop.chisq = FALSE,
           chisq = TRUE,
           sresid = TRUE)
```

Does it meet assumptions and why or why not?: 

Interpret the results:


### Flashlight and zombie status

H0: 

HA: 

#### Make a prediction about what you will find based on exploratory analysis: 

```{r flashlightZombieBarGraph}
# flashlight & zombie status (see page 289)
zombies %>% 
  group_by(flashlight, zombie) %>% 
  count() %>% 
  group_by(flashlight) %>% 
  mutate(perc.flashlight = 100*n/sum(n)) %>% 
  ggplot(aes(x = flashlight, y = perc.flashlight, fill = zombie)) +
  geom_col(position = "dodge") +
  theme_minimal() +
  labs(y = "Percent of zombies within flashlight group",
       x = "Has flashlight")
```

Prediction: 

#### Conduct the analysis

```{r flashlightZombieChiSq}
# flashlight and zombie status
CrossTable(y = zombies$zombie,
           x = zombies$flashlight,
           expected = TRUE,
           prop.c = FALSE,
           prop.t = FALSE,
           prop.chisq = FALSE,
           chisq = TRUE,
           sresid = TRUE)
```

Does it meet assumptions and why or why not?: 


Interpret the results:

### First aid kit and zombie status

H0: 

HA: 

#### Make a prediction about what you will find based on exploratory analysis: 

```{r firstAidZombieExplore}
# first aid kit & zombie status (see page 289)

```

Prediction: 

#### Conduct the analysis

```{r firstAidZombieChiSq}
# first aid kit and zombie status
```

Does it meet assumptions and why or why not?: 


Interpret the results:

## Independent samples t-tests

Independent samples t-tests are used to compare the means of a continuous variable across two groups (PAGE: 343, 331)

In determining the characteristics associated with being a zombie, which is a categorical variable, which of the variables in this data set would be examined using a t-test (delete the ones that are not):

* **age:** age in years
* **waterPerPerson:** gallons of water per person

## What are the assumptions for an independent samples t-test? (PAGE: 365)

* Normal distribution within each group
* Independent observations
* Equal variances for the two groups
* One continuous variable and two independent groups

## Conducting t-tests

### Age and zombie status

H0: There is no relationship between age and zombie status. There is no difference in mean age for zombies and non zombies.

HA: There is a difference in mean age for zombies and non zombies.

#### Make a prediction about what you will find based on exploratory analysis: 

```{r densityAgeZombie}
# age & zombie status (see page 346)
zombies %>% 
  ggplot(aes(x = age, fill = zombie)) +
  geom_density(alpha = .8) +
  theme_minimal() +
  labs(x = "Distribution of age",
       y = "Probability density")
```

Prediction: 

#### Conduct the analysis

```{r ttestAgeZombie}
# age and zombie status
t.test(formula = zombies$age ~ zombies$zombie,
       var.equal = TRUE)
```

Does it meet assumptions and why or why not?: 



```{r densityAgeZombieFacets}
# distribution
zombies %>% 
  ggplot(aes(x = age)) +
  geom_density() +
  facet_grid(cols = vars(zombie)) +
  theme_minimal()
```

```{r leveneAgeZombie}
# equal variances 
car::leveneTest(y = age ~ zombie, data = zombies)
```

Interpret the results: 

```{r mannwhitUageZombie}
# mann-whitney U
wilcox.test(formula = zombies$age ~ zombies$zombie,
            paired = FALSE)
```


[disclaimer: since we didn't meet assumptions, we'd actually do a mann-whitney U test instead]

### Water per person and zombie status

H0: The mean gallons of water per person is the same between zombies and non zombies

HA: The mean gallons of water per person is not the same for zombies and non zombies

#### Make a prediction about what you will find based on exploratory analysis: 

```{r boxplotWaterZombie}
# water per person & zombie status (see page 346)
zombies %>% 
  ggplot(aes(y = waterPerPerson, x = zombie)) +
  geom_boxplot() +
  theme_minimal() +
  labs(x = "Zombie status",
       y = "Gallons of water per person")
```

Prediction: 

```{r ttestWaterZombie}
# water per person and zombie status
t.test(formula = zombies$waterPerPerson ~ zombies$zombie,
       var.equal = TRUE)
```

Does it meet assumptions and why or why not?: 



```{r graphWaterZombie}
# distribution
zombies %>% 
  ggplot(aes(x = waterPerPerson)) +
  geom_density() +
  facet_grid(cols = vars(zombie)) +
  theme_minimal()
```

```{r leveneTestWaterZombie}
#equal variances
car::leveneTest(y = waterPerPerson ~ zombie, data = zombies)
```


Interpret the results:


[disclaimer: we did not meet assumptions so would usually choose a different analysis strategy; categorize water per person and use chi-squared]

## Overall interpretation

What is associated with being a zombie? 


## If there is time...a code shortcut 

```{r shortcut}
library(package = "tableone")

# create table for zombie data
# with zombie (outcome variable) as the strata
# table now shows p-values for bivariate tests
# chi-squared for categorical+categorical
# t-test for continuous+binary
# Mann-Whitney U test for non-normal continuous+binary
zombieTable <- CreateTableOne(data = zombies, 
                              vars = c("age", "gender",
                                       "rurality", "food",
                                       "radio", "flashlight",
                                       "firstaid", "waterPerPerson"),
                              strata = "zombie")
print(zombieTable, showAllLevels = TRUE,
      nonnormal = c("age", "waterPerPerson"))
``` 


## The End