---
title: "Week 4 workshop"
output: html_document
---

Goal: Use the BRFSS 2021 data to review code and concepts from Chapters 1-3.

Research goal: Exploratory analysis for a study that examines poor physical health days (PHYSHLTH) and insurance status (HLTHPLN1) as predictors of whether someone got a colonoscopy (HADSIGM4).

Activities:

TOGETHER:

1) Import data
2) Examine imported data 
3) Use the codebook to learn about the variables (https://www.cdc.gov/brfss/annual_data/2021/pdf/codebook21_llcp-v2-508.pdf)
4) Select the variables of interest into a smaller data frame
5) Recode the variables so that "don't know" and "refused" are treated as missing 
6) Drop any observations that are missing data on any of the 3 variables of interest 
7) Add labels to factor variables and do any other cleaning needed
8) Visualize each variable with a well-formatted plot

GROUP ACTIVITY: Work with a small group to find and complete the "TASK" sections of this research report. Pick a person from your group to keep track of your work and report out to the larger group. 

TASK 1: 

Find each of the following variables in the codebook and indicate the meaning of the variable and how it was measured for each one. Use HADSIGM4 as a model for the others.

1) HADSIGM4
Question: Colonoscopy and sigmoidoscopy are exams to check for colon cancer. Have you ever had either of these exams? 
Responses: 1 = Yes, 2 = No, 7 = Don't know/Not sure, 9 = Refused, blank is missing
Notes: Not asked if participant AGE is less than 45
Data type: categorical 

2) _HLTHPLN
Question: 
Responses:
Notes:
Data type: 

3) PHYSHLTH 
Question: 
Responses:
Notes:
Data type: 


```{r openPackages}
# open the packages need to import (haven) and manage (tidyverse) the data
library(package = "tidyverse")
library(package = "haven")
library(package = "table1")
```

```{r openData}
# import data (box on pages 59-60)
temp <- tempfile(fileext = ".zip")
download.file(url = "https://www.cdc.gov/brfss/annual_data/2021/files/LLCP2021XPT.zip", 
              destfile = temp)
brfss <- read_xpt(file = temp)

# uncomment the next line and run it to save a local copy
# this can be useful if it takes a long time to download the
# web version of the data
# write.csv(brfss, "brfss.csv")
```

```{r summarizeData}
# Examine imported data
# this code is an alternative to clicking on the name of the data
# in the environment window
View(x = brfss)
```

```{r cleanData}
# select and manage variables
# codebook: https://www.cdc.gov/brfss/annual_data/2021/pdf/codebook21_llcp-v2-508.pdf
brfss.little <- brfss %>% 
  select(PHYSHLTH, HADSIGM4, `_HLTHPLN`) %>% 
  mutate(PHYSHLTH = as.numeric(PHYSHLTH)) %>% 
  mutate(PHYSHLTH = na_if(x = PHYSHLTH, y = 77)) %>%
  mutate(PHYSHLTH = na_if(x = PHYSHLTH, y = 99)) %>%
  mutate(PHYSHLTH = recode(PHYSHLTH, `88` = 0)) %>%
  mutate(HADSIGM4 = na_if(x = HADSIGM4, y = 7)) %>%
  mutate(HADSIGM4 = na_if(x = HADSIGM4, y = 9)) %>%
  mutate(`_HLTHPLN` = na_if(x = `_HLTHPLN`, y = 7)) %>%
  mutate(`_HLTHPLN` = na_if(x = `_HLTHPLN`, y = 9)) %>%
  drop_na() %>% 
  mutate(HADSIGM4 = recode_factor(HADSIGM4,
                                `1` = "Yes",
                                `2` = "No")) %>% 
  mutate(`_HLTHPLN` = recode_factor(`_HLTHPLN`,
                                `1` = "Yes",
                                `2` = "No")) %>% 
  rename(daysBadPhysHealth = PHYSHLTH) %>% 
  rename(hasInsurance = `_HLTHPLN`) %>% 
  rename(hadColonoscopy = HADSIGM4)

# save to a new file
# uncomment to use this code
# write.csv(brfss.little, "brfsslittle.csv")

# load the new file
# uncomment to use this code
# brfssClean <- read.csv("brfsslittle.csv")
```

TASK 2

Finish this list to describe how each of the functions was used in the data management:

* The select() function was used to...
* The as.numeric() function was used to...
* The na_if() function was used to...
* The recode() function was used to...
* The recode_factor() function was used to...
* The rename() function was used to...
* The drop_na() function was used to...

```{r summarizeCleanedData}
# summarize the managed data set
summary(object = brfss.little)
```

```{r summaryTable}
# make a table of the managed data set
# the dot is used to signify that all variables in the data
# should be in the table
table1(~ .,
       data = brfss.little)
```

TASK 3

Add bullet points interpreting the descriptive statistics in your table:

* 
*
*


TASK 4 

Add the appropriate geom layer to the code to choose an appropriate graph type for each of the variables. Different data types will have different types of graphs (e.g., bar graph, histogram, density plot).

Add a layer with a title and better labels for the x and y axes to each graph.

Make any other formatting changes you'd like (e.g., add a theme, add color).

Interpret what you see in each graph in one sentence. 

```{r insuranceGraph}
# visualize hasInsurance
brfss.little %>% 
  ggplot(aes(x = hasInsurance)) +
  # add the appropriate geom here 
```

Interpretation:

```{r colonoscopyGraph}
# visualize hadColonoscopy
brfss.little %>% 
  ggplot(aes(x = hadColonoscopy)) +
  # add the appropriate geom here
```

Interpretation: 

```{r physHealthGraph}
# visualize daysBadPhysHealth 
brfss.little %>% 
  ggplot(aes(x = daysBadPhysHealth)) +
  # add the appropriate geom here

```

Interpretation: 

Which descriptive statistics would be best for the days of poor health variable?



-----------------------

HACKER TASK (EXTRA/OPTIONAL)

For the hasInsurance and hadColonoscopy graphs, change the statistic shown on the y-axis to a percentage rather than a frequency.

TRY SOMETHING NEW (EXTRA/OPTIONAL)

Code chunk options can be used in R Markdown to format your document so that you can create fully reproducible documents.

* Knit the file and review it
* Go back and add to the {r} to the second code chunk so it looks like this: {r include = FALSE}
* Knit the file again to see what changes
* What does adding include = FALSE do?
* For the very first code chunk, add to the {r} so it looks like this: {r message = FALSE}
* What does adding message = FALSE do? 

Explore the code chunk options for R Markdown here: https://rmarkdown.rstudio.com/lesson-3.html

Try adding options to your code chunks and knitting it until you have a clean knitted document.
