---
title: "Week 4 workshop: Chapters 1-3 Review"
output: html_document
---

Goal: Use the BRFSS 2014 data to review code and concepts from Chapters 1-3.

Research goal: Exploratory analysis for a study that examines poor physical health days (PHYSHLTH) and insurance status (HLTHPLN1) as predictors of whether someone got a mammogram (HADMAM).

Activities:

1) Import data
2) Examine imported data 
3) Use the codebook to learn about the variables (https://www.cdc.gov/brfss/annual_data/2014/pdf/CODEBOOK14_LLCP.pdf)
4) Select the variables of interest into a smaller data frame
5) Recode the variables so that "don't know" and "refused" are treated as missing 
6) Drop any observations that are missing data on any of the 3 variables of interest 
7) Add labels to factor variables and do any other cleaning needed
8) Visualize each variable with a well-formatted plot

GROUP ACTIVITY: Work with your colleagues to find and complete the "TASK" sections of this research report. Choose one person from your group to report out to the larger group. 


```{r openData}
# open the packages need to import (haven) and manage (tidyverse) the data
library(package = "tidyverse")
library(package = "haven")

# 1) Import data (box on pages 59-60)
temp <- tempfile(fileext = ".zip")
download.file(url = "http://www.cdc.gov/brfss/annual_data/2014/files/LLCP2014XPT.zip", destfile = temp)
brfss <- read_xpt(file = temp)

# write.csv(brfss, "brfss.csv")
```

```{r summarizeData}
# 2) Examine imported data
summary(object = brfss)
```

```{r cleanData}
# 3-7) select and manage variables
# codebook: https://www.cdc.gov/brfss/annual_data/2014/pdf/CODEBOOK14_LLCP.pdf
brfss.little <- brfss %>% 
  select(PHYSHLTH, HADMAM, HLTHPLN1) %>% 
  mutate(PHYSHLTH = as.numeric(PHYSHLTH)) %>% 
  mutate(PHYSHLTH = na_if(x = PHYSHLTH, y = 77)) %>%
  mutate(PHYSHLTH = na_if(x = PHYSHLTH, y = 99)) %>%
  mutate(PHYSHLTH = recode(PHYSHLTH, `88` = 0)) %>%
  mutate(HADMAM = na_if(x = HADMAM, y = 7)) %>%
  mutate(HADMAM = na_if(x = HADMAM, y = 9)) %>%
  mutate(HLTHPLN1 = na_if(x = HLTHPLN1, y = 7)) %>%
  mutate(HLTHPLN1 = na_if(x = HLTHPLN1, y = 9)) %>%
  drop_na() %>% 
  mutate(HADMAM = recode_factor(HADMAM,
                                `1` = "Yes",
                                `2` = "No")) %>% 
  mutate(HLTHPLN1 = recode_factor(HLTHPLN1,
                                `1` = "Yes",
                                `2` = "No")) %>% 
  rename(daysBadPhysHealth = PHYSHLTH) %>% 
  rename(hasInsurance = HLTHPLN1) %>% 
  rename(hadMammogram = HADMAM)
  
# write.csv(brfss.little, "brfsslittle.csv")
```

TASK 1

Finish this list to describe what was done in the data management:

* The select() function chose the three variables from the larger data set
* The as.numeric() function...
* The na_if() function...
* The recode_factor() function...
* The drop_na() function...
* The rename() function...

```{r summarizeCleanedData}
# summarize the managed data set
summary(object = brfss.little)
```

TASK 2

Finish this list interpreting the summary information:

* Of the 249,002 survey participants, 231,280 have health insurance and 17,722 do not have health insurance
* 
* 


```{r insuranceGraph}
# 8) visualize hasInsurance
brfss.little %>% 
  ggplot(aes(x = hasInsurance, fill = hasInsurance)) +
  geom_bar() +
  theme_minimal() +
  labs(x = "Participant has health insurance",
       y = "Number of participants") +
  scale_fill_manual(values = c("#7463AC", "deeppink"), guide = "none")
# note that guide = FALSE is outdated and now guide = "none" is used instead
```

Interpretation: 

TASK 3 

Create and interpret appropriate and well-formatted visualizations for the other two variables.

```{r physHealthGraph}
# 8) visualize daysBadPhysHealth 

```

Interpretation: 

```{r mammogramGraph}
# 8) visualize hadMammogram

```

Interpretation: 


HACKER TASK (EXTRA/OPTIONAL)

For the hasInsurance and hadMammogram graphs, change the statistic shown on the y-axis to a percentage rather than a frequency.

TRY SOMETHING NEW (EXTRA/OPTIONAL)

Code chunk options can be used in R Markdown to format your document so that you can create fully reproducible documents.

* Knit the file and review it
* Go back and add to the {r} to the second code chunk so it looks like this: {r include = FALSE}
* Knit the file again to see what changes
* What does adding include = FALSE do?
* For the very first code chunk, add to the {r} so it looks like this: {r message = FALSE}
* What does adding message = FALSE do? 

Explore the code chunk options for R Markdown here: https://rmarkdown.rstudio.com/lesson-3.html

Try adding options to your code chunks and knitting it until you have a clean knitted document.
