---
title: "R You Ready? An Introduction to R"
subtitle: "Week 1 Workshop"
output: html_document
---

## How to follow along and participate

* Install the `tidyverse`, `table1`, and `haven` packages 

    + Use the `Tools` menu and choose the `Install packages...` menu choice 

    + Type the names of the packages, `tidyverse, table1, haven`, in the dialog box that opens and click Install

* Click on the **tiny** green triangle on the top right corner of the gray shaded area (this is a code chunk!) that contains this: 

```{r open-packages}
library(package = "tidyverse")
library(package = "table1")
library(package = "haven")
library(package = "sjlabelled")
library(package = "janitor")
```

* The arrow will run all the code in that "chunk" of code

* Throughout this workshop you will be able to run the code from the demonstration by clicking the arrow on the top right in the corresponding code chunk

* In between code chunks, you can type as you would in a Word document if you'd like to take notes


## Importing data into R


```{r import-data}
# use read_spss function to read the spss file 
globalAttitudes2023 <- read_sav(file = "Pew Research Center Global Attitudes Spring 2023 Dataset.sav") %>% 
  clean_names()
```

## Choosing variables

```{r select-variables}
# use select to choose variables
globalAttitudes2023Rec <- globalAttitudes2023 %>% 
  select(country, age, sex, abortion_legal, gay_marriage)
```

## Check your work

```{r summarize-data}
summary(object = globalAttitudes2023Rec)
```

## Choosing variables & observations

```{r choose-variables-observations}
globalAttitudes2023Rec <- globalAttitudes2023 %>% 
  select(country, age, sex, abortion_legal, gay_marriage) %>% 
  filter(age > 25)
```

## Check your work

```{r check-work}
summary(object = globalAttitudes2023Rec)
```

## Managing data types in R

* Commonly used data types in R:

    + factor: variable with categories (categorical data) 
    
    + numeric: variable with values over some continuum (continuous data) 
    
    + integer: whole number with values over some continuum (often treated as continuous)
    
    + character: used for words or other verbatim information (e.g., ZIP codes, names) 
    
* Several of the variables are the wrong data type! Which ones?
    
* The `mutate()` function is useful for changing variable types or values or really anything 

* We can add this function into our data management chain

## Recoding variables

```{r recode-age-gender}
# recode age and sex
globalAttitudes2023Rec <- globalAttitudes2023 %>% 
  select(country, age, sex, abortion_legal, gay_marriage) %>% 
  filter(age > 25) %>% 
  mutate(age = as.numeric(age)) %>% 
  mutate(sex = as_label(sex))
```

## Check your work

```{r check-gender-age-recode}
summary(object = globalAttitudes2023Rec)
```

## Recoding more variables

```{r recode-abortion}
# recode race-ethnicity
globalAttitudes2023Rec <- globalAttitudes2023 %>% 
  select(country, age, sex, abortion_legal, gay_marriage) %>% 
  filter(age > 25) %>% 
  mutate(age = as.numeric(age)) %>% 
  mutate(sex = as_label(sex)) %>% 
  mutate(abortion_legal = as_label(abortion_legal))
```

## Check your work

```{r check-abortion-recode}
summary(object = globalAttitudes2023Rec)
```
## Recode country

```{r recode-country}
# recode country
globalAttitudes2023Rec <- globalAttitudes2023 %>% 
  select(country, age, sex, abortion_legal, gay_marriage) %>% 
  filter(age > 25) %>% 
  mutate(age = as.numeric(age)) %>% 
  mutate(sex = as_label(sex)) %>% 
  mutate(abortion_legal = as_label(abortion_legal)) %>% 
  mutate(country = as_label(country))
```

## Check your work

```{r check-country-recode}
summary(object = globalAttitudes2023Rec)
```

## Correcting the age variable

* Check the codebook to see that the values of 98 and 99 are "Don't know" and "Refused" and do not represent people who are actually 98 and 99 years old

* Use the replace() function to replace these values with NA, which is "not available" or missing in R

```{r fix-age-missing}
# recode country
globalAttitudes2023Rec <- globalAttitudes2023 %>% 
  select(country, age, sex, abortion_legal, gay_marriage) %>% 
  filter(age > 25) %>% 
  mutate(age = as.numeric(age)) %>% 
  mutate(age = replace(age, age > 97, NA)) %>% 
  mutate(sex = as_label(sex)) %>% 
  mutate(abortion_legal = as_label(abortion_legal)) %>% 
  mutate(country = as_label(country))
```


## You Try It!

* Finish the recoding yourself by adding the gay_marriage variable to the data cleaning

* When finished adding this code, run the `summary()` function to see if it works

```{r you-try-it-recode}
# add the gay marriage variable to the end of this data management code
# remember to use the pipe to move from one mutate to the next
globalAttitudes2023Rec <- globalAttitudes2023 %>% 
  select(country, age, sex, abortion_legal, gay_marriage) %>% 
  filter(age > 25) %>% 
  mutate(age = as.numeric(age)) %>% 
  mutate(age = replace(age, age > 97, NA)) %>% 
  mutate(sex = as_label(sex)) %>% 
  mutate(abortion_legal = as_label(abortion_legal)) %>% 
  mutate(country = as_label(country)) 
```

```{r check-your-recode-you-try-it}
# check your work
summary(object = globalAttitudes2023Rec)
```

## Basic table

```{r basic-table-of-data}
# basic table with 4 variables
table1(~ sex + abortion_legal + gay_marriage + age,
       data = globalAttitudes2023Rec)
```


## Fixing labels

```{r format-table}
label(globalAttitudes2023Rec$sex) <- "Participant sex"
label(globalAttitudes2023Rec$age) <- "Age in years"
label(globalAttitudes2023Rec$abortion_legal) <- "Abortion should be"
label(globalAttitudes2023Rec$gay_marriage) <- "Allow gay marriage"

units(globalAttitudes2023Rec$age) <- "years"
```


## Fancier table

```{r fancy-table}
table1(~ sex + abortion_legal + gay_marriage + age,
       data = globalAttitudes2023Rec,
       render.continuous = c(. = "Mean (SD)"),
       caption = "Characteristics and Attitudes of Global Survey Participants, 2023 (n = 21,672)")
```

## Stratified table

```{r stratified-table}
table1(~ abortion_legal + gay_marriage + age | sex,
       data = globalAttitudes2023Rec,
       render.continuous = c(. = "Mean (SD)"),
       caption = "Characteristics and Attitudes of Global Survey Participants by Sex, 2023 (n = 21,672)")
```


## Basic bar chart

```{r bar-chart}
globalAttitudes2023Rec %>%
  ggplot(aes(x = gay_marriage)) +
  geom_bar() 
```

## The end