---
title: "My Life in Months"
author: "Jenine Harris"
date: "8/5/2023"
output: html_document
---

# STEP 1: TYPE IN DATA

Type notes in here to remember what we did or any details you think are important.

```{r chunk1-data}
# this is a code chunk where you write R code
# if you add a hashtag to start a line it is a comment like this

# ADD SOME DATA
# the eras of my life
# CHANGE THIS CODE TO YOUR ERAS
era <- c("1971 Eugene, Oregon, USA", 
         "1988 Corvallis, Oregon, USA", 
         "1989 Eugene, Oregon, USA",
         "1992 Avignon, France", 
         "1992 Eugene, Oregon, USA",
         "1993 Washington DC, USA", 
         "1994 Eugene, Oregon, USA", 
         "2000 Bend, Oregon, USA", 
         "2001 St. Louis, Missouri, USA")

# the months I spent in each era (same number and order as eras)
# CHANGE THIS CODE TO YOUR MONTHS
months <- c(213, 
            21, 
            18,
            4, 
            12,
            16, 
            67, 
            12, 
            270) 
```

# STEP 2: COMBINE THE DATA INTO A DATA FRAME

```{r chunk2-combine}
# STEP 2: Make the graph
# open the tidyverse package (install first through Tools menu)
library(package = "tidyverse")

# the bar plot needs a data frame rather than a table
# combine the vectors into a data frame
my_life_df <- data.frame(era, months)
my_life_df$era <- factor(my_life_df$era, levels = my_life_df$era)
```

# STEP 3: MAKE THE GRAPH

Try replacing "deeppink" with a color you like, like "red" or "blue" or find a color on the https://colorbrewer2.org/ website and add it where it currently says deeppink. 

```{r chunk3-graph}
# make the plot
my_life_df %>% 
  ggplot(aes(x = era, y = months)) +
  geom_col(fill = "deeppink") +
  coord_flip() +
  theme_minimal() +
  labs(x = "Where I lived", y = "Number of months")
```

# STEP 4 (optional): MAKE IT A WAFFLE!

Waffle plots are a better alternative to a pie chart for showing parts of a whole. Unfortunately, the waffle making R package is a little buggy, so it can be a little limited and might not work on your computer. But, let's try it!

```{r chunk4-waffle-data}
# open the waffle package (install first through Tools menu)
library(package = "waffle")

# make a new vector with no more than 8 eras 
# fewer is better for this type of graph
# CHANGE THESE NUMBERS TO YOUR NUMBERS
my_life_waffle <- c(310,
                    21,
                    4,
                    16,
                    12,
                    270) 
# CHANGE THE WORDS INSIDE THE QUOTATION MARKS TO YOUR WORDS
names(my_life_waffle) <- c("Eugene, Oregon, USA",
                           "Corvallis, Oregon, USA",
                           "Avignon, France",
                           "Washington DC, USA",
                           "Bend, Oregon, USA",
                           "St. Louis, Missouri, USA")
```

```{r chunk5-waffle-plot}
# make the plot
waffle(my_life_waffle / 4, rows = 6,
       legend_pos = "bottom")
# once it runs, try replacing the 4 and 6 with different 
# numbers and see what happens
# the 4 divides the number of months by 4 so each square is worth 4
# months, the 6 is the number of rows in your waffle 
# try replacing "bottom" with "top" and see what happens
```

# Resources for doing more

Bar graph options: https://ggplot2.tidyverse.org/reference/geom_bar.html

Colors: https://colorbrewer2.org/